Self-tunable approximated explicit MPC: Heat exchanger implementation

### 2023-01-19 paper/elsarticle/main.tex
paper: results part1

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

%% main text
\section{Introduction}
\label{sec:intro}

\section{Preliminaries}
\label{}

\section{Methodology}
\label{}

\section{Results and discussion}
\label{}

\section{Conclusion}
\label{}

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num} 
%%  \bibliography{<your bibdatabase>}

@@

\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

\bibitem{}

\end{thebibliography}
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{lineno}
\usepackage[colorlinks,citecolor=black,linkcolor=black,urlcolor=black,bookmarks,bookmarksnumbered]{hyperref}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{amsmath}

@@

%% main text
\section{Introduction}
\label{sec:introduction}

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Explicit model predictive control}
\label{sec:eMPC}
TODO: remark, ze sme si vedomi aj delta-u formulacie, ale vedie na zlozitejsie eMPC

\subsection{Tunable explicit model predictive control}
\label{sec:tunable}

\subsection{Self-tunable explicit model predictive control}
\label{sec:self_tunable}

\section{Methodology}
\label{sec:methodology}

\subsection{Self-tunable technique for systems with asymmetric behavior}
\label{sec:self_tunable_new}

TODO: spomenut, ze to je vhodne celkovo aj pre klasicke online MPC

\section{Results and discussion}
\label{sec:results}

In this section, the results of the proposed tuning method are presented. The plant on which the control was implemented is a laboratory liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Fig.~\ref{fig:HE}. The cold feed as well as heating medium are transported to the heat exchanger by two peristaltic pumps. The flow rate of the feed is constant, while the aim is to track the reference value of its temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger and the manipulated variable is the voltage $U$ corresponding to the power of the heating medium pump. The voltage is set in percentage, while it is constrained from 20\% to 100\%. As the heat exchange is a nonlinear and non-symmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy.

The matrices of the linear state-space model of the plant, discretized with sampling time $T_\mathrm{s}$ = 1\,s, are

\begin{subequations}
	\label{eq:model_A_B} 
	\begin{eqnarray}
		A&=&\begin{bmatrix}
			0.839
		\end{bmatrix}, \\
		B&=&\begin{bmatrix}
			0.039
		\end{bmatrix}, \\
		C&=&\begin{bmatrix}
			1
		\end{bmatrix}, \\
		D&=&\begin{bmatrix}
			0
		\end{bmatrix}.
		\end{eqnarray}
\end{subequations}

The constraints are considered in the terms of manipulated variable, i.e.

\begin{eqnarray}
\label{eq:u_const}
	-15\% \le u \le 65\%.
\end{eqnarray}

Note, the variable $u$ represents the manipulated variable in the deviation form. The values of feed temperature and voltage of the heating medium pump corresponding to zero steady state are respectively $T^\mathrm{s}$~=~35 $^{\circ}\mathrm{C}$ and $U^\mathrm{s}$~=~35\%.

The penalty matrices of the problem TODO:ref were systematically tuned and the corresponding control was implemented on the laboratory heat exchanger for every MPC setup. The aim was to determine, which penalty matrix will be the tuned one. Based on observations, the most significant effect on the control trajectories had tuning the penalty matrix $Q_\mathrm{y}$, while still preserving a satisfactory control performance, i.e., without steady-state control error and significant oscillations around the setpoint. The boundary values of matrix $Q_\mathrm{y}$ were tuned as $Q_\mathrm{y, L}$ = 100 and $Q_\mathrm{y, U}$ = 1000. The built-in integrator was penalized with $Q_\mathrm{x2}$ = 1 and the manipulated variable with $R$ = 10. The prediction horizon $N$ was 20 steps long. The control results for both control setups are compared in Fig.~\ref{fig:CV_boundaries} for the controlled variable, and in Fig.~\ref{fig:MV_boundaries} for the manipulated variable.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/HE}
		\caption[Heat exchanger Armfield Process Plant Trainer PCT23]{Laboratory heat exchanger Armfield Process Plant Trainer PCT23: feed pump\,(1), heating medium pump\,(2), feed tanks\,(3), heater for heating medium\,(4), heat exchanger\,(5).}
		\label{fig:HE}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/CV_boundaries}
		\caption{Controlled variable trajectory for two boundary controllers. The solid lines represent the controlled temperature and the dashed line represents the reference value.}
		\label{fig:CV_boundaries}
	\end{center}
\end{figure}

The trajectories in Fig.~\ref{fig:CV_boundaries} show the non-symmetric nature of controlling the process of heat exchange mainly when observing the overshoots and undershoots. When applying the control associated with the lower bound $Q_\mathrm{y, L}$, significant undershoots are present when tracking the reference downwards, i.e., when the reference change is negative. On the contrary, when implementing the controller associated with $Q_\mathrm{y, U}$, the undershoots are negligible, but significant overshoots can be seen when tracking the reference upwards.

These observations form the base for the strategy of self-tuning the penalty matrix $Q_\mathrm{y}$. The strategy follows the ideas summarized in Section~\ref{sec:self_tunable_new}.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/MV_boundaries}
		\caption{Manipulated variable trajectory for two boundary controllers. The solid lines represent the voltage and the dashed lines represent the constraints.}
		\label{fig:MV_boundaries}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/CV}
		\caption{Controlled variable trajectory for two boundary controllers and the tuned one. The solid lines represent the controlled temperature and the dashed line represents the reference value.}
		\label{fig:CV}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/MV}
		\caption{Manipulated variable trajectory for two boundary controllers and the tuned one. The solid lines represent the voltage and the dashed lines represent the constraints.}
		\label{fig:MV}
	\end{center}
\end{figure}

\begin{table}[h!]
	\begin{center}
		\caption{Control performance criteria.}
		\label{tab:control_performance}
		\begin{tabular}{c|c|c|c|c|c} 
			Reference step change & $Q_\mathrm{y}$ & ISE [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s] & $V$\,[l] \\
			\hline
			\multirow{2}{*}{1} & 1000 & 714 & 33.50 & 16.5 & \textbf{2.12} \\
			    & 100 & 867 & 16.65 & 12.5 & 2.36 \\ 
			    & tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5} & 2.38 \\ 
			\hline
			\multirow{2}{*}{2} & 1000 & 365 & 47.20 & \textbf{5} & \textbf{2.49} \\
			    & 100 & 606 & 23.25 & 26.5 & 3.19 \\ 
			    & tuned & \textbf{248} & \textbf{19.13} & 9.5 & 3.35 \\ 
			\hline
		    \multirow{2}{*}{3} & 1000 & 245 & \textbf{18.92} & \textbf{6.5} & \textbf{2.00} \\
				& 100 & 398 & 79.64 & 31 & \textbf{2.00} \\ 
				& tuned & \textbf{186} & 24.59 & \textbf{6.5} & 2.10 \\ 
			\hline
			\multirow{2}{*}{4} & 1000 & 1024 & 18.43 & 22.5 & 0.94 \\
				& 100 & 1402 & 41.87 & 90 & \textbf{0.93} \\ 
				& tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5} & 1.10  
		\end{tabular}
	\end{center}
\end{table}

\section{Conclusion}
\label{sec:conclusion}

@@

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%

\bibliographystyle{elsarticle-num} 
\bibliography{references}

@@

%\begin{thebibliography}{00}
%
%%% \bibitem{label}
%%% Text of bibliographic item
%
%\bibitem{}
%
%\end{thebibliography}

\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.
